Keyword and Collocation analysis on India-related text of *Natural History*

### 2023-08-24 paper_quarto.qmd
0824 wording refine

knitr::write_bib(
  c("base", "mclm", "ggplot2", "tidyverse", "leaflet", "downlit"),
  here("packages.bib")
)
```

@@

[^1]: This study is a pilot study of my master thesis, so the description of the research question is connected with that of this [paper](https://raw.githack.com/lizaodawn/NH_thesis/main/NHthesis_structure.html#sec-introduction).

Pliny the Elder's *Natural History* is widely recognized as the earliest encyclopedia in the world. Written around AD 77, the work is thematically divided into 37 books, covering a diverse range of subjects including astronomy, geography, zoology, botany, medicine, and more. Driven by a pioneering attempt to catalogue human knowledge from his era, Pliny made a comprehensive reference to a wide range of Greek and Roman scholarships, and interwove his own literary interpretation or comments to the narratives in *Natural Hisrory*.

*Natural History* stands as a seminal work in the history of encyclopedias, serving as a gateway to understanding the knowledge and worldview of ancient Rome. Beyond its role as a compendium of diverse subjects, the text provides a captivating glimpse into the knowledge, cultural, and merchandise exchange [@dodd2021] between the Roman Empire and other regions while shedding a light on the spatial perception and imagination of Pliny and the scholarly references of his time [@plinyth2011].

As pointed out by @beagon2011, differentiating from his predecessors, Pliny showed a "terrestrial curiosity" in Natural History, emphasising recognition of the physical, material world. Drawing from the long-established topographical and ethnographic traditions, Pliny connects volumes dedicated to geography (books 3-6) with broader elements, activities, and cultural, historical, and societal contexts [@roller2022], exemplified in his portrayal of exotic plants, communities' habitats, imperial expeditions, and trade products. In this regard, geographical names mentioned in the text plays a pivotal role in distributing information, knowledge, and discussions throughout *Natural History*.

*Natural History* is originally in Latin. For this study, an English translation by Henry T. Riley (1816-1878) and John Bostock (1773-1846), first published in 1855, is utilized. The translated text is obtained in a digitised version from the [TOPOSText project](https://topostext.org/the-project), having been sourced from the Perseus Project and governed by a Creative Commons Attribution-Share-Alike 3.0 U.S. License.

In addition to the digitised textual content of *Natural History*, the place names mentioned in the text are annotated with their corresponding coordinates on TOPOXText. The annotations are available in an HTML format, which were sparsed and restructured into a dataset of geographical-related text in *Natural History* with [Beautiful Soup](https://www.crummy.com/software/BeautifulSoup/bs4/doc/) library in Python[^2]. The dataset structure, corpus scraping procedure and and corpus extention will be further elaborated in Corpus Description section (@sec-corpus).

[^2]: The sparsing [code](https://github.com/lizaodawn/zhuang/blob/papertry/Natural%20History_corpus.ipynb) is uploaded to the same repository of the project.

As a starting point, @fig-geoname reveals the Top 20 distinct geographical names mentioned within the text.

```{r}
#| label: fig-geoname
#| fig-cap: Bar chart for top 20 distinct place names mentioned in *Natural History*
#| fig-cap-location: top
#| warning: false
#| echo: false

@@

# Create a bar chart using ggplot2
top_20_summary_data <- summary_data %>%
  top_n(20, Count) %>%
  ungroup()

@@

An intriguing observation from this analysis, as depicted in @fig-map, is that India seems prominently mentioned in the work despite being outside the Mediterranean.

```{r}
#| label: fig-map
#| fig-cap: Mapping of geographical names mentioned in Natural History
#| fig-cap-location: top
#| warning: false
#| echo: false

@@

Building upon this observation, the study focus is narrowed down to India-related text in *Natural History*, aiming to delve deeper into the themes associated with the Indian region referred in the text by employing collocation analysis and keyword analysis methods within the framework of corpus linguistics.

The analysis is conducted in R environment [@R-base], and the main anlysing methods are applied through functions from [mclm](https://masterclm.github.io/mclm/) [@R-mclm] package. In addition, [tidyverse](https://www.tidyverse.org/) [@R-tidyverse] and [leaflet](https://rstudio.github.io/leaflet/) [@R-leaflet] are used for data manipulation and visualisation.

# Methodology

To investigate the prominent themes in the India-related content discussed in *Natural History*, methods of collocation analysis and keyword analysis are employed for this study. Both of the analyses rely on the `mclm::assoc_scores` objects of the [mclm](https://github.com/masterclm/mclm) package, which return frequencies and association scores of a certain event in a target context and in a reference context.

Both of the methods will be examined by three association measurements, namely absolute frequency, PMI and $G^2$ signed, to discuss with the context in *Natural History* and the potential theme it indicates.

PMI measures the association strength between two words by comparing their observed co-occurrence frequency with their expected co-occurrence frequency by chance. It highlights words that appear together more frequently than expected. High PMI values indicates strong associations. High $G^2$ values suggest that the co-occurrence is unlikely to occur by chance.

$G^2$ is a statistical test that measures the independence or association between two categorical variables (words in this case). It compares the observed co-occurrence frequencies with the expected frequencies under the assumption of independence. Higher $G^2$ values suggest a stronger association.

@@

Collocation analysis is a technique employed in corpus linguistics to identify words that often appear together in a language. This technique is rooted in the notion that particular words tend to appear together more often than would occur by chance. In the context of this study, lists of the highly possible co-occuring word with "India" in the text of *Natural History* can shed a light on the potential themes and topics about India discussed in the work.

Combining the two methods, and integrating with a close reading of the text, this study aims to discuss what about "India" is discussed in *Natural History*.

@@

Keyword analysis is a statistical method used in corpus linguistics to identify words that occur more frequently in a target corpus than in a reference corpus. Defined with different associate measurements, the "keyness" of words are calculated and indicated which words are statistically more frequent appeared in the target corpus comparing to that of the reference corpus.

In this study, the *target corpus* includes all paragraphs in *Natural History* mentioning Indian regions, and the **reference corpus** contains all other paragraphs in the work. Lists of 20 words with the highest score in different association measurements are discussed with the textual context. As indicated by the measurements, the listed words tend to be more frequently occur in the context pertaining India than other parts of the work, hence they seem to be the potential keywords about India in *Natural History*.

@@

The digitised text is avalible in paragraphs with geographical annotations on TOPOSText website as two parts, [book 1-11](https://topostext.org/work/148) and [book 12-37](https://topostext.org/work/153), both of which are in HTML format. Extracting all the "place" class in the HTML files, a dataset containing all annotated place names in the work, the URI and corresponding coordinates of the place, the book/chapter/paragraph number where it is mentioned, as well as the coordinating textual content are structured as a dataset of geographic-related text in *Natural History*. This dataset is used for an initial exploration of the prominent mentioned places in the text.

Since the research focus is narrowed down to India-related text after the initial exploration, a dataset including all Indian place names mentioned in the text is extracted from the geographical-related text dataset. With a reference to *Barrington Atlas of the Greek and Roman World* (@talbert2000, @talbert2000-1), the approximate coordinates defining the Indian region during the time of *Natural Hisoty* are as follows[^3]:

@@

Each distinct paragraph in the India-related text dataset is extracted and stored in TXT format as separate files within a corpus folder, comprising a total of 146 files as the **target corpus** for India-related text in *Natural History*. The file names contain information about the book, chapter, and paragraph numbers.

And the full text of the entire work, segregating in paragraphs, are scraped and stored in TXT format and excluded the files in **target corpus** to serve as a **reference corpus**.

@@

@@

# Data Analysis

## Collocating word to "India" in *Natural History*

In addition to examining the keywords in the India-related texts, it is also valuable to explore the collocates that appear alongside the word "India" when mentioned in the book.

To analyse the textual collocating bi-gram of word "India", the `mclm::text_cooc()` function was employed, with "india" designated as the target item for the analysis. Since the texual collocation of "India" is examinated, the span of the target word is set as the "textual units", to take the broader context into consideration.

And following the same criteria as the previous keyword analysis, the collocation scores were filtered of having both PMI and G_signed scores of 2 or higher. By employing this approach, the prominent collocates of "india" were identified, providing a tangible understanding of the prevailing discourse and potentially reflecting Pliny the Elder's perspective on the region.

@@

top_scores_colloc <- scores_colloc %>% 
  filter(PMI >= 2 & G_signed >= 2)
```

@tbl-coltable_Gsigned and @tbl-coltable_PMI display the top 20 collocates associated with the word "india" in the entire text, respectively sorted by G_signed and PMI scores. As the returned $G^2$ score helps gauge the strength of evidence regarding the independence between the two words [@methods], thereby facilitating the initial sorting.

@@

```
```{r}

@@

The frequency lists for the **target corpus** (India-related texts) and **reference corpus** (whole text) are generated respectively, and the association scores are calculated through the `mclm::assoc_scores()` function. The association score is filtered based on the criteria of both PMI and $G^2$ scores, requiring a minimum score of 2. This criterion ensures that the selected words have a probability at least four times higher than the overall data, along with statistically significant supporting evidence [@methods], allowing to identify the most statistically significant keywords in India-related texts when compared to the reference corpus.

@@

top_scores_kw <- scores_kw %>% 
  filter(PMI >= 2 & G_signed >= 2)

@@

The following tables present the top 20 keywords in India-related texts as compared to the entire texts in *Natural History*. They are sorted based on their PMI (@tbl-kwtable_PMI), Absolute frequencies(@tbl-kwtable_absolute_freqencies) and G_signed scores (@tbl-kwtable_Gsigned), respectively. The initial sorting was carried out based on the PMI scores (@tbl-kwtable_PMI), since the it serves as a crucial indicator for measuring the keyness of words.

>>>>

knitr::write_bib(
  c("base", "mclm", "tidyverse", "leaflet", "kableExtra", "downlit"),
  here("packages.bib")
)
```

@@

[^1]: The present paper is a pilot study of my master thesis, *Mapping India in Pliny the Elder's Natural History*, thus the description of the research question is connected with the content in this [thesis](https://raw.githack.com/lizaodawn/NH_thesis/main/NHthesis_structure.html#sec-introduction).

Pliny the Elder's *Natural History* is widely recognized as the earliest encyclopedia in the world. Written around AD 77, the work is thematically divided into 37 books, covering a diverse range of subjects including astronomy, geography, zoology, botany, medicine, and more. Driven by a pioneering attempt to catalog human knowledge from his era, Pliny made a comprehensive reference to a wide range of Greek and Roman scholarships, and interwove his own literary interpretation or comments to the narratives in *Natural History*.

*Natural History* stands as a seminal work in the history of encyclopedias, serving as a gateway to understanding the knowledge and worldview of ancient Rome. Beyond its role as a compendium of diverse subjects, the text provides a captivating glimpse into the knowledge, cultural, and merchandise exchange [@dodd2021] between the Roman Empire and other regions [@plinyth2011].

As pointed out by @beagon2011, differentiating from his predecessors, Pliny showed a "terrestrial curiosity" in *Natural History*, emphasizing recognition of the physical, material world. Drawing from the long-established topographical and ethnographic traditions, Pliny connects volumes dedicated to geography (books 3-6) with broader natural elements, human activities, and cultural, historical, and societal contexts [@roller2022], exemplified in his description of plants, tribe habitats, imperial expeditions, and trade products. In this regard, this study aims to analyse the text of *Natural History* in a spatial scope, as geographical names mentioned in the text plays a pivotal role in distributing information, knowledge, and discussions throughout the work.

*Natural History* is originally in Latin. An English translation by Henry T. Riley (1816-1878) and John Bostock (1773-1846), first published in 1855, is utilized for this study. The translated text is obtained in a digitized version from the [TOPOSText project](https://topostext.org/the-project), having been sourced from the Perseus Project and governed by a Creative Commons Attribution-Share-Alike 3.0 U.S. License.

In addition to the digitized textual content of *Natural History*, the place names mentioned in the text are annotated with their corresponding coordinates on TOPOXText. Since the annotations are available in an HTML format, they are scraped and restructured into a dataset of geographical-related text in *Natural History* with [Beautiful Soup](https://www.crummy.com/software/BeautifulSoup/bs4/doc/) library in Python[^2]. The dataset structure, corpus scraping procedure and and corpus extension will be further explained in Corpus Description section (@sec-corpus).

[^2]: The scraping [code](https://github.com/lizaodawn/zhuang/blob/papertry/Natural%20History_corpus.ipynb) is uploaded to the same repository of the project.

As a starting point of this study, @fig-top20geoname reveals the 20 most frequently mentioned geographical names within the text.

```{r}
#| label: fig-top20geoname
#| fig-cap: Top 20 most frequently mentioned place names in *Natural History*

@@

top_20_summary_data <- summary_data %>%
  top_n(20, Count) %>%
  ungroup()

@@

Additionally, @fig-geonamemap depicted the times of each place name found mentioned in the work on a map. The individual places are represented by distinct dots, and its occurring times in the text are represented by the size of the dots. The larger the dot size, the more frequent it is appeared in the context.

```{r}
#| label: fig-geonamemap
#| fig-cap: Mapping of geographical names mentioned in *Natural History*
#| fig-cap-location: top
#| warning: false
#| echo: false

@@

An intriguing observation from these figures is that India seems being prominently mentioned in the work despite locating outside the Mediterranean.

Building upon this observation, the study focus is narrowed down to India-related text in *Natural History*. By delving deeper into the themes associated with the Indian region referred in the text, this study aims to examine the portrayal of India by Pliny the Elder, which potentially shed a light on the spatial perception and imagination of the representative writers and scholars of ancient Rome.

# Methodology {#sec-method}

To investigate the prominent themes in the India-related content discussed in *Natural History*, the study generally employs methods of collocation analysis and keyword analysis in corpus linguistics, and is conducted in R environment [@R-base]. The discussions are based on the `mclm::assoc_scores` objects returned from [mclm](https://masterclm.github.io/mclm/) [@R-mclm] package. In addition, [tidyverse](https://www.tidyverse.org/) [@R-tidyverse] and [leaflet](https://rstudio.github.io/leaflet/) [@R-leaflet] packages are used for data manipulation and visualization.

In the Data Analysis section (@sec-data_analysis), both of the collocation and keyword analysis are examined by three association measures, namely absolute frequency, PMI and signed $G^2$ for discussing the potential theme it inferred integrating with a close reading of the context.

Absolute frequency refers to the simple count of each single type (distinct word) observed in the text. In the case of collocation analysis, it shows the time a type collocates with the target word. While in the case of keyword analysis, it shows the time a type occurs in the target corpus. As a basic counting measure, absolute frequency provide an overview of the word distribution with limited insights.

PMI (Pointwise Mutual Information) is one of effect size measures for association. It compares the observed co-occurring frequency of two outcomes to the probability of either of them occurring independently. It can be applied to observe word association, and the association between words and context categories as well [@bestvater2022]. As the PMI value increases by one, the probability becomes twice as high. High PMI values indicates strong associations.

$G^2$ refers to the log-likelihood ratio for measuring the strength of statistical evidence of the observasion [@blume2017]. In the context of this study, it indicates how significant the association between the two factors (either for two words or for keywords associating with a target corpus) is. Specificly, the signed $G^2$ score returned by `mclm::assoc_scores` shows the strength of evidence for attraction, i.e. co-occurrence. High signed $G^2$ values suggest that the co-occurrence is unlikely to occur by chance.

Since a PMI score higher than two indicates a four times higher probability of association, and a signed $G^2$ score higher than 3.84 indicates a significant association at a 95% confidence level [@montes2022], a filter of "PMI \>= 2 & G_signed \>= 3.84" is applied for the returned objects of `mclm::assoc_scores` for both of the methods, in order to gain strong and significant association outputs for further discussion.

@@

Collocation analysis is a technique employed in corpus linguistics to identify words that often appear together in a given text. This technique is rooted in the notion that particular words tend to appear together more often than they would occur by chance.

In this study, the textual collocating word with "India" in *Natural History* is examined. Lists of highly frequent co-occurring word with "India" in the context are obtained and sorted by scores of different measurements. The potential theme of India-related content in the work can be discussed based on the output of textual collocation of word "India".

@@

Keyword analysis is a statistical method used in corpus linguistics to identify words that occur more frequently in a target corpus than in a reference corpus. The words showing higher association with the target corpus comparing to its frequency in the reference corpus are seen as the "keyword" of the target corpus.

In this study, the **target corpus** includes all paragraphs in *Natural History* mentioning place names in Indian regions, and the **reference corpus** contains all other paragraphs in the work. Lists of words with relatively higher frequency in the target corpus are obtained and sorted by scores of different measurements. The "keywords" found in India-related text in *Natural History* provides more descriptive information about the topics and elements related to "India" in Pliny's narrative.

Combining the two methods, and integrating with a close reading of the text, this study aims to gain insights about how "India" is discussed and portrayed in *Natural History*.

@@

The digitized text is available in paragraphs with geographical annotations on TOPOSText website as two parts, [book 1-11](https://topostext.org/work/148) and [book 12-37](https://topostext.org/work/153), both of which are in HTML format. Extracting all "place" class in the HTML files, information of all annotated place names in the work, the URI and corresponding coordinates of the place, the book/chapter/paragraph number where it is mentioned, as well as the coordinating textual content is structured as a dataset of geographical-related text in *Natural History*. This dataset is used for an initial exploration of the prominent mentioned places in the text.

The initial obeservation shown in @fig-top20geoname and @fig-geonamemap leads to a focus to the India-related text for this study. With a reference to *Barrington Atlas of the Greek and Roman World* (@talbert2000; @talbert2000-1), the approximate geographical coordinates defining the Indian region during the time of *Natural Hisoty* are as follows[^3]:

@@

@@

Each distinct paragraph in the India-related text dataset is extracted and stored in TXT format as separate files within a corpus folder, comprising a total of `length(fname_indiatext)` files as the **target corpus**. The file names contain information about the book, chapter, and paragraph number of the text.

And the full text of the entire work, separated in paragraphs, are also scraped and stored in TXT format. Excluding those in **target corpus**, the other part of the work, containing `length(fname_wholetext_filtered)` files, serve as a **reference corpus**.

@@

# Data Analysis {#sec-data_analysis}

## Textual collocating word with "India" in *Natural History*

To analyse the textual collocating word with "India", the `mclm::text_cooc()` function was employed, with "India" designated as the target item for the analysis. Since the textual collocation of "India" is examined, the span of the target word is set as the "textual units", to take the broader context into consideration.

As mentioned in the Methodology Section (@sec-method), the collocation scores were filtered for those having PMI scores higher than 2 and signed $G^2$ scores higher than 3.84. By employing this approach, the prominent collocates of "India" were identified, providing a tangible understanding of the prevailing discourse and potentially reflecting Pliny the Elder's perspective on the region.

@@

top_scores_colloc <- scores_colloc %>% 
  filter(PMI >= 2 & G_signed >= 3.84)
```

@tbl-coltable_Gsigned and @tbl-coltable_PMI display the top 20 collocates associated with the word "india" in the entire text, respectively sorted by G_signed and PMI scores. As the returned $G^2$ score helps gauge the strength of evidence regarding the independence between the two words [@montes2022], thereby facilitating the initial sorting.

@@

```{r}

@@

The frequency lists for the **target corpus** (India-related texts) and **reference corpus** (whole text) are generated respectively, and the association scores are calculated through the `mclm::assoc_scores()` function. The association score is filtered based on the criteria of both PMI and $G^2$ scores, requiring a minimum score of 2. This criterion ensures that the selected words have a probability at least four times higher than the overall data, along with statistically significant supporting evidence [@montes2022], allowing to identify the most statistically significant keywords in India-related texts when compared to the reference corpus.

@@

top_scores_kw <- scores_kw %>% 
  filter(PMI >= 2 & G_signed >= 3.84)

@@

The following tables present the top 20 keywords in India-related texts as compared to the entire texts in *Natural History*. They are sorted based on their PMI (@tbl-kwtable_PMI), Absolute frequencies (@tbl-kwtable_absolute_freqencies) and G_signed scores (@tbl-kwtable_Gsigned), respectively. The initial sorting was carried out based on the PMI scores (@tbl-kwtable_PMI), since the it serves as a crucial indicator for measuring the keyness of words.
